

One of the challenging problems in earth sciences is to incorporate the vast quantities of data that are constantly being collected world-wide into dynamical models for these systems, and is called the problem of data assimilation (DA). DA is a crucial ingredient for making meaningful real time predictions such as weather forecasts, hurricane tracking, and possibly even climate predictions. \cite{asch2016data, carrassi2018data} The Bayesian formulation of DA naturally leads to the problem of nonlinear filtering, which studies the conditional distribution, called the \emph{filter} or the \emph{posterior} distribution, of the state at any time conditioned on observations up to that time. \cite{ApteH07, law2015data, reich2015probabilistic}

A natural question is about the stability of the filter with respect to the initial condition, which is the probability distribution of the state at the initial time. This question has been studied extensively, e.g., \cite{bishop2017stability, chigansky2006stability}, but mostly in the context of stochastic systems. In many applications in the earth sciences, the models used are deterministic and only a few results about filter stability are known, see {\cite{reddy2019asymptotic,  reddy2021stability}} and the references therein. The main focus of this chapter is to illustrate numerically the stability of two commonly used filtering algorithms, namely the particle filter and the ensemble Kalman filter.

The main contributions of this chapter are as follows. The most commonly used method for assessing the filter stability is using the root-mean square error (RMSE) as the distance between the filter mean and the true trajectory. But such a measure at best implies stability of the mean. In this chapter, we explicitly calculate distances between filtering distributions starting with different initial conditions, thus assessing stability directly (see the definition~\ref{def-stab--numerical-fs}). For this study, we apply the recently developed algorithms~\cite{feydy2019interpolating, genevay2019entropy, thibault2021overrelaxed} for calculating optimal transport distances between distributions relevant to the filtering problem.

% https://www.overleaf.com/project/60adddaf1944021514796207
% https://www.overleaf.com/project/60adddaf1944021514796207
